Evaluation of probabilistic photometric redshift estimation approaches for \boldblue{ \textsc{The Rubin Observatory Legacy Survey of Space and Time} (LSST)}

### 2020-03-13 paper/tex/appendix.tex
change bootstraps to 44k and address point 17

Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
The horizontal lines show the the result of a bootstrap resampling of the training set using 30,000 samples for \trainz, representing a conservative idealized limit on expected performance for a modest-sized representative training set of galaxies, as mentioned in Section~\ref{sec:pitqq}.
The AD bootstrap statistic is elevated due to its sensitivity to the tails of distributions.
The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, whereas \eazy, \lephare, \metaphor, and \skynet\ underperform; \bpz, \gpz, and \tpz\ are within a factor of two of the conservative limit for all statistics.
It is unsurprising that \cmnn\ scores well, as with a nearly complete and representative training set choosing neighbouring points in colour/magnitude space to construct an estimator should lead to excellent agreement in the final $\hat{N}(z)$.

>>>>

Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
The horizontal lines show the the result of a bootstrap resampling of the training set using 30,000 samples for \trainz, representing a conservative \sout{idealized limit on expected performance for}\boldblue{error estimate assuming} a modest-sized representative training set of galaxies, as mentioned in Section~\ref{sec:pitqq}.
The AD bootstrap statistic is elevated due to its sensitivity to the tails of distributions.
The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, whereas \eazy, \lephare, \metaphor, and \skynet\ underperform; \bpz, \gpz, and \tpz\ are within a factor of two of the conservative limit for all statistics.
It is unsurprising that \cmnn\ scores well, as with a nearly complete and representative training set choosing neighbouring points in colour/magnitude space to construct an estimator should lead to excellent agreement in the final $\hat{N}(z)$.

### 2020-03-13 paper/tex/results.tex
embiggen labels on Fig 2
change bootstraps to 44k and address point 17

The gray shaded band marks the $2\sigma$ variance in PIT values found using the \trainz\ algorithm with a bootstrap resampling of the training set and a sample size of 30,000 galaxies, representing a very conservative estimate of the representative training sample size estimated as being required for direct \pz\ calibration \citep{Newman:2015}, and thus an approximate minimal error significance compared to ideal performance.
The existence of deviations in the PIT histograms outside of this gray shaded uncertainty range show that significant biases are present for some codes.

\begin{figure*}
\centering
\includegraphics[width=0.74\textwidth]{fig/PITANDQQplot_12codes_withsigmaband_crop.jpg}
\caption{The QQ plot (red) and PIT histogram (blue) of the \pzpdf\ codes (panels) along with the ideal QQ (black dashed diagonal) and ideal PIT (gray horizontal) curves, as well as a difference plot for the QQ difference from the ideal diagonal (lower inset).
The gray shaded region indicates the $2 \sigma$ range from a bootstrap resampling of the training set with a size of 30,000 galaxies using \trainz.
The twelve codes exhibit varying degrees of four deviations from perfection: an overabundance of PIT values at the centre of the distribution indicate a catalogue of overly broad \pzpdf s, an excess of PIT values at the extrema indicates a catalogue of overly narrow \pzpdf s, catastrophic outliers manifest as overabundances at PIT values of 0 and 1, and asymmetry indicates systematic bias, a form of model misspecification.
Values in excess of the $2\sigma$ shaded region show that for some codes these errors will be significant given expected training sample sizes.}
\label{fig:pitqq}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
\caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
There is generally good agreement between these statistics, with differences corresponding to the codes with outstanding catastrophic outlier rates, a reflection of the differences in how each statistic weights the tails of the distribution.
Horizontal lines indicate the level of uncertainty found by bootstrapping a training set sample of 30,000 galaxies using \trainz; none of the codes reach this conservative ideal floor in expected uncertainty.}
\label{fig:pit_stats}
\end{figure*}

Figure~\ref{fig:pit_stats} highlights the relative values of the KS, CvM, and AD test statistics calculated by comparing the PIT distribution and a uniform distribution $U(0, 1)$.
\metaphor\ and \lephare\ perform well under the AD but poorly under the KS and CvM due to their high catastrophic outlier rates.
\annz\ and \flexzboost\  are the top scorers under these metrics of the PIT distribution.
\annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with PIT values that more closely match the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.
Similar to what was done for the PIT histograms in Figure~\ref{fig:pitqq}, we create bootstrap training samples of 30,000 galaxies for use with \trainz\ in order to estimate a conservative statistical floor that we would expect in real data.
No code reaches this idealized floor, indicating that all codes suffer some degradation from the ideal when employing their implicit priors, though \annz, \flexzboost, and \gpz\ are within a factor of two.

>>>>

The gray shaded band marks the $2\sigma$ variance in PIT values found using the \trainz\ algorithm with a bootstrap resampling of the training set and a sample size of \sout{30}\boldblue{44},000 galaxies, representing a \sout{very} conservative estimate of the \boldblue{expected errors due to the training sample size} \sout{representative training sample size estimated as being required for direct \pz\ calibration  }, and thus an approximate minimal error significance compared to ideal performance.
The existence of deviations in the PIT histograms outside of this gray shaded uncertainty range show that significant biases are present for some codes.

\begin{figure*}
\centering
\includegraphics[width=0.74\textwidth]{fig/PITANDQQplot_12codes_withsigmaband_biglabels_crop.jpg}
\caption{The QQ plot (red) and PIT histogram (blue) of the \pzpdf\ codes (panels) along with the ideal QQ (black dashed diagonal) and ideal PIT (gray horizontal) curves, as well as a difference plot for the QQ difference from the ideal diagonal (lower inset).
The gray shaded region indicates the $2 \sigma$ range from a bootstrap resampling of the training set with a size of \sout{30}\boldblue{44},000 galaxies using \trainz.
The twelve codes exhibit varying degrees of four deviations from perfection: an overabundance of PIT values at the centre of the distribution indicate a catalogue of overly broad \pzpdf s, an excess of PIT values at the extrema indicates a catalogue of overly narrow \pzpdf s, catastrophic outliers manifest as overabundances at PIT values of 0 and 1, and asymmetry indicates systematic bias, a form of model misspecification.
Values in excess of the $2\sigma$ shaded region show that for some codes these errors will be significant given expected training sample sizes.}
\label{fig:pitqq}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_44k.jpg}
\caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.
There is generally good agreement between these statistics, with differences corresponding to the codes with outstanding catastrophic outlier rates, a reflection of the differences in how each statistic weights the tails of the distribution.
Horizontal lines indicate the level of uncertainty found by bootstrapping a training set sample of \sout{30}\boldblue{44},000 galaxies using \trainz; none of the codes reach this conservative ideal floor in expected uncertainty.}
\label{fig:pit_stats}
\end{figure*}

Figure~\ref{fig:pit_stats} highlights the relative values of the KS, CvM, and AD test statistics calculated by comparing the PIT distribution and a uniform distribution $U(0, 1)$.
\metaphor\ and \lephare\ perform well under the AD but poorly under the KS and CvM due to their high catastrophic outlier rates.
\annz\ and \flexzboost\  are the top scorers under these metrics of the PIT distribution.
\annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with PIT values that more closely match the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.
Similar to what was done for the PIT histograms in Figure~\ref{fig:pitqq}, we create bootstrap training samples of \sout{30}\boldblue{44},000 galaxies for use with \trainz\ in order to estimate a conservative statistical floor that we would expect in real data.
No code reaches this idealized floor, indicating that all codes suffer some degradation from the ideal when employing their implicit priors, though \annz, \flexzboost, and \gpz\ are within \boldblue{roughly} a factor of two.
